Visual diagnostics for constrained optimisation with application to guided tours

### 2021-01-02 paper/paper.tex
last go-through before the vis-diag section

The goal of continuous optimisation is to find the best solution within
the space of all feasible solutions where typically the best solution is
decided by an objective function. Broadly speaking optimization can be
unconstrained or constrained \citep{kelley1999iterative}. The
unconstrained problem can be formulated as a minimization (or
maximization) problem such as \(\min_{x} f(x)\) where
\(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with
certain properties defined in an \(L^p\) space. In this case, solutions
rely on gradient descent or ascent methods. In the constrained
optimization problem additional restrictions are introduced via a set of
functions that can be convex or non-convex:
\(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and
hence the problem can be written as \(\min_{x} f(x)\) \emph{subject to}
\(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex
optimization methods including linear and quadratic programming can be
used.

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational} which is an
exploratory data analysis tool that is defined to detect
\emph{interesting structures} or features in high-dimensional data
through a set of lower-dimensional projections that cover the entire
high dimensional space using interpolation methods called tours
\citep{cook2008grand}. The target of the optimisation is to identify the
most \emph{interesting} low-dimensional views of the data given by a
corresponding projection matrix. The most \emph{interesting} structures
are formally defined by a function of projections, called index function
which is optimized to uncover the most reveling structures in a high
dimensional space \citep{cook1993projection}.

The optimization challenges encountered in the projection pursuit guided
tour problem are common to those of optimization in general. Examples of
those include the existence of multiple maxima (local and global), the
trade off between computational burden and proximity to the maxima,
dealing with noisy objective functions that might be non smooth and non
diferentiable \citep{jones1998efficient}. Those are not unique to this
context and therefore the visualization tools and optimization methods
presented in this paper can be easily apply to any other optimization
problems.

@@

Optimization problems are ubiquitous in many areas of study. While in
some cases analytical solutions can be found, the majority of problems
rely on numerical methods to find the optimal solution. These numerical
methods follow iterative approaches that aim at finding the optimum by
progressively improving the current solution until a desirable accuracy
is achieved. Although that principle seems uncomplicated, optimization
methods need to deal with many challenges such as the possible existence
of multiple maxima (local and global), the trade-off between desirable
accuracy and computational burden, as well as dealing with noisy
objective functions and possible constraints. In addition, the
optimization results might depend on the algorithm starting values,
affecting the consistency of results.

Our interest is on constrained optimization
\citep{bertsekas2014constrained} as defined in the introduction section,
and assuming it is not possible to find a solution to the problem in the
way of a closed-form. That is, the problem consists of finding the
minimum or maximum of a function \(f \in L^p\) in the constrained
\(\mathbb{A}\) space.

Optimization methods can be divided into various classes, such as global
optimisation \citep{kelley1999iterative, fletcher2013practical}, convex
optimisation \citep{boyd2004convex} or stochastic optimisation
\citep{nocedal2006numerical}. A large class of methods utilises the
gradient information in the objective function with the most notable one
being the gradient ascent (descent) method. Although gradient
optimization methods are popular, they rely on the availability of the
objective function derivatives and on the complexity of the constraints.
Because of that, we consider another class of method in this paper:
derivative-free methods. Derivative-free methods have been developed
over the years, where the emphasis is on finding, in most cases, a near
optimal solution. Examples of those include response surface methodology
\citep{box1951experimental}, stochastic approximation
\citep{robbins1951stochastic}, random search \citep{fu2015handbook} and
heuristic methods \citep{sorensen2013metaheuristics}. Later, we will
present a simulated annealing optimisation algorithm, which belongs to
the class of random search methods, for optimisation with the guided
tour.

A common search scheme utilised by both derivative-free methods and
gradient methods is line search. In line search methods, users are
required to provide an initial estimate \(x_{1}\) and at each iteration,
a search direction \(S_k\) and a step size \(\alpha_k\) are generated.
Then one moves on to the next point following
\(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
desire convergence is reached. While gradient-based methods choose the
search direction by the gradient, derivative-free methods determine the
search direction through local information of the objective function.
The choice of step size also needs considerations, as inadequate step
sizes might prevent the optimisation method to converge to an extremum.
An ideal step size can be chosen via finding the value of
\(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
respect to \(\alpha_k\) at each iteration.

@@

A number of index functions have been proposed in the literature to
detect different data structures, including Legendre index
\citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, natural Hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} and PDA
index \citep{lee2010projection}.

As a general visualisation method, a tour produces animations of high
dimensional data via rotations between low dimension planes. Different
tour types choose these planes differently, for example, a grand tour
\citep{cook2008grand} selects the planes randomly and a manual tour
\citep{cook1997manual} gradually phases in and out one variable, to
understand the contribution of that variable in the projection. Guided
tour, the main interest of this paper, chooses planes with the aid of
projection pursuit. Given a random start, projection pursuit iteratively
finds bases with higher index values and the guided tour constructs the
geodesic interpolation between these planes to form a tour path.

Mathematical details of the geodesic interpolation can be found in
\citet{buja2005computational}. Figure \ref{fig:tour-path} shows a sketch
of the tour path. The blue frames are produced by the projection pursuit
optimisation algorithm, and the white frames interpolate between them.
The tour method has been implemented in the R package \CRANpkg{tourr}
\citep{tourr}.

@@

pa The optimisation problem in the tour context is stated as follows:
\ldots. .Given a randomly generated starting basis \(\mathbf{A}_1\),
projection pursuit finds the final projection basis \(\mathbf{A}_T\)
that satisfies the following optimisation problem:

@@

Below we introduce three possible algorithms: \texttt{search\_better},
\texttt{search\_better\_random}, and \texttt{search\_geodesic}. The
first two are derivative free methods that sample candidate bases in the
neighbourhood whilst \texttt{search\_geodesic} is an analogue of
gradient ascent on the projection basis space.

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $\alpha$, $l_{\max}$} 
    \Output{$\mathbf{A}_{l}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    Compute $I_{l}  = f(\mathbf{A}_{l})$\;
    \If{$I_{l} > I_{\text{cur}}$}{
      \KwRet{$\mathbf{A}_{l}$} \;
      }
    $l = l + 1$\;
  }
  \caption{random search}
  \label{random-search}
\end{algorithm}

\texttt{search\_better} is a random search device that samples a
candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
basis \(\mathbf{A}_{\text{cur}}\) by
\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
where \(\alpha\) controls the radius of the sampling neighbourhood and
\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
then orthogonalised to ensure the orthonormal constraint is fulfilled.
When a basis is found with index value higher than the current basis
\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
for guided tour to construct an interpolation path. The next iteration
of search begins after adjusting \(\alpha\) by a cooling parameter:
\(\alpha_{j+1} = \alpha_j * \text{cooling}\). The termination condition
is when the maximum number of iteration \(l_{\max}\) is reached. The
algorithm of \texttt{search\_better} is summarised in Algorithm
\ref{random-search}. A slightly different cooling scheme has been
proposed by \citet{posse1995projection} to include a halving parameter
\(c\). Rather than reducing the radius of the searching neighbourhood,
\(\alpha\), at each iteration, Posse's design only adjust \(\alpha\) if
the last search takes more than \(c\) times to find an accepted basis to
avoid the searching space being reduced too fast.

\begin{algorithm}
\SetAlgoLined
    Compute $I_{l} = f(\mathbf{A}_{l})$ and $T(l) = \frac{T_0}{\log(l + 1)}$\;
      \eIf{$I_{l} > I_{\text{cur}}$}{
        \KwRet{$\mathbf{A}_{l}$} \;
      }{
        Compute $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
        Draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
        \If{$P > U$}{
           \KwRet{$\mathbf{A}_{l}$} \;
        }
      }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

Simulated annealing (\texttt{search\_better\_random})
\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
same sampling process as \texttt{search\_better} but allows a
probabilistic acceptance of a basis with lower index value based on the
annealing \(T(l)\). Given an initial \(T_0\), the temperature at
iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
candidate basis fails to have an index value larger than the current
basis, simulated annealing gives it a second chance to be accepted with
probability
\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
where \(I_{(\cdot)}\) denotes the index value of a given basis. This
implementation allows the algorithm to jump out of a local maximum and
enables a more holistic search of the whole parameter space. This
feature is particularly useful when local maxima are present. The
algorithm can be written as replacing line 5-8 of Algorithm
\ref{random-search} with Algorithm \ref{simulated_annealing}.

\begin{algorithm}
\SetAlgoLined
\SetKwInOut{Input}{input}
  \SetKwInOut{Output}{output}
    \Input{$\mathbf{A}_{\text{cur}}$, $f$, $l_{\max}$, $n = 5$, $\delta$}
    \Output{$\mathbf{A}_{**}$}
  initialisation\;
  Set $l = 1$\;
  \While{$l < l_{\max}$}{
    Generate $2n$ bases in a small neighbourhood, $\delta$, of $\mathbf{A}_{\text{cur}}$ and ensure orthogonality \;
    Find the one with the largest index value: $\mathbf{A}_{*}$\;
    Construct the geodesic $\mathcal{G}$ from $\mathbf{A}_{\text{cur}}$ to $\mathbf{A}_{*}$\;
    Optimise the index value on the geodesic $\mathcal{G}$ over a 90 degree window to produce the optima $\mathbf{A}_{**}$  \;
    Compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
      \If{$p_{\text{diff}} > 0.001$}{
         \KwRet{$\mathbf{A}_{**}$} \;
      }
    $l = l + 1$\;
  }
  \caption{search geodesic}
  \label{search-geodesic}
\end{algorithm}

\citet{cook1995grand} used a gradient ascent algorithm on the space of
the projection bases. In gradient ascent, one first finds the direction
for improvement via computing the gradient information. In
\texttt{search\_geodesic}, \(2n\) bases are first generated in a tiny
neighbourhood of the current basis, controlled by the neighbourhood
parameter \(\delta\). A geodesic is then constructed using the current
basis and the one in \(2n\) bases with the highest index value. If the
neighbourhood parameter \(\delta\) is tiny, the geodesic constructed is
an analogue of the gradient information in the curved space and works as
the searching direction. The next step in gradient ascent is to conduct
a line search to find the best improvement along the gradient direction
and in \texttt{search\_geodesic}, this is replaced by optimising the
index value along the geodesic direction over an 90 degree angle from
\(-\pi/4\) to \(\pi/4\). The optima \(\mathbf{A}_{**}\) is returned for
the current iteration if it meets the termination condition on
percentage improvement. The procedure will also terminate if
\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
steps in geodesic search.

@@

\address{%
Nicolas Langrené\\
Data61 CSIRO\\
Locked Bag 38004 Docklands VIC 8012 Australia\\
}
\href{mailto:Nicolas.Langrene@data61.csiro.au}{\nolinkurl{Nicolas.Langrene@data61.csiro.au}}

>>>>

The goal of continuous optimisation is to find the best solution within
the space of all feasible solutions where typically the best solution is
decided by an objective function. Broadly speaking, optimization can be
unconstrained or constrained \citep{kelley1999iterative}. The
unconstrained problem can be formulated as a minimization (or
maximization) problem such as \(\min_{x} f(x)\) where
\(f:\mathbb{R}^n \rightarrow \mathbb{R}\) is an objective function with
certain properties defined in an \(L^p\) space. In this case, solutions
rely on gradient descent or ascent methods. In the constrained
optimization problem additional restrictions are introduced via a set of
functions that can be convex or non-convex:
\(g_i:\mathbb{R}^n \rightarrow \mathbb{R}\) for \(i = 1, \ldots k\) and
hence the problem can be written as \(\min_{x} f(x)\) \emph{subject to}
\(g_i(x) \leq 0\). Here methods such as Langrange multipliers and convex
optimization methods including linear and quadratic programming can be
used.

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational} which is an
exploratory data analysis tool that is defined to detect
\emph{interesting structures} or features in high-dimensional data
through a set of lower-dimensional projections that cover the entire
high dimensional space using interpolation methods called tours
\citep{cook2008grand}. The target of the optimisation is to identify the
most \emph{interesting} low-dimensional views of the data given by a
corresponding projection matrix. The most \emph{interesting} structures
are formally defined by a function of projections, called index function
which is optimized to uncover the most revelling structures in a high
dimensional space \citep{cook1993projection}.

The optimization challenges encountered in the projection pursuit guided
tour problem are common to those of optimization in general. Examples of
those include the existence of multiple maxima (local and global), the
trade off between computational burden and proximity to the maxima,
dealing with noisy objective functions that might be non smooth and non
differentiable \citep{jones1998efficient}. Those are not unique to this
context and therefore the visualization tools and optimization methods
presented in this paper can be easily applied to any other optimization
problems.

@@

Optimization problems are ubiquitous in many areas of study. While in
some cases analytical solutions can be found, the majority of problems
rely on numerical methods to find the optimal solution. These numerical
methods follow iterative approaches that aim at finding the optimum by
progressively improving the current solution until a desirable accuracy
is achieved. Although this principle seems uncomplicated, a number of
challenges arise such as the possible existence of multiple maxima
(local and global), constraints and noisy objective function, and the
trade-off between desirable accuracy and computational burden. In
addition, the optimization results might depend on the algorithm
starting values, affecting the consistency of results.

Optimization methods can be divided into various classes, such as global
optimisation \citep{kelley1999iterative, fletcher2013practical}, convex
optimisation \citep{boyd2004convex} or stochastic optimisation
\citep{nocedal2006numerical}. Our interest is on constrained
optimization \citep{bertsekas2014constrained} as defined in the
introduction section, and assuming it is not possible to find a solution
to the problem in the way of a closed-form. That is, the problem
consists of finding the minimum or maximum of a function \(f \in L^p\)
in the constrained \(\mathbb{A}\) space.

A large class of methods utilises the gradient information of the
objective function to perform the optimisation iterations, with the most
notable one being the gradient ascent (descent) method. Although
gradient optimization methods are popular, they rely on the availability
of the objective function derivatives and on the complexity of the
constraints. Derivative-free methods, which do not rely on the knowledge
of the gradient, are more generally applicable. Derivative-free methods
have been developed over the years, where the emphasis is on finding, in
most cases, a near optimal solution. Examples of those include response
surface methodology \citep{box1951experimental}, stochastic
approximation \citep{robbins1951stochastic}, random search
\citep{fu2015handbook} and heuristic methods
\citep{sorensen2013metaheuristics}. Later, we will present a simulated
annealing optimisation algorithm, which belongs to the class of random
search methods, for optimisation with the guided tour.

A common search scheme utilised by both derivative-free methods and
gradient methods is line search. In line search methods, users are
required to provide an initial estimate \(x_{1}\) and, at each
iteration, a search direction \(S_k\) and a step size \(\alpha_k\) are
generated. Then one moves on to the next point following
\(x_{k+1} = x_k + \alpha_kS_k\) and the process is repeated until the
desire convergence is reached. While gradient-based methods choose the
search direction by the gradient, derivative-free methods uses local
information of the objective function to determine the search direction.
The choice of step size also needs considerations, as inadequate step
sizes might prevent the optimisation method to converge to an optimum.
An ideal step size can be chosen via finding the value of
\(\alpha_k \in \mathbb{R}\) that maximises \(f(x_k + \alpha_kS_k)\) with
respect to \(\alpha_k\) at each iteration.

@@

A number of index functions have been proposed in the literature to
detect different data structures, including Legendre index
\citep{friedman1974projection}, Hermite index
\citep{hall1989polynomial}, natural Hermite index
\citep{cook1993projection}, chi-square index
\citep{posse1995projection}, LDA index \citep{lee2005projection} for
supervised classification problems, and PDA index
\citep{lee2010projection}, which is an extension of the LDA index.

As a general visualisation method, a tour produces animations of high
dimensional data via rotations between low dimension planes. Different
tour types choose these planes differently, for example, a grand tour
\citep{cook2008grand} selects the planes randomly to provide a general
overview and a manual tour \citep{cook1997manual} gradually phases in
and out one variable, to understand the contribution of that variable in
the projection. Guided tour, the main interest of this paper, chooses
planes with the aid of projection pursuit to gradually reveal the most
interesting projection in the low dimension space. Given a random start,
projection pursuit iteratively finds bases with higher index values and
the guided tour constructs the geodesic interpolation between these
planes to form a tour path. Intuitively, Figure \ref{fig:tour-path}
shows a sketch of the tour path where the blue frames are produced by
the projection pursuit optimisation algorithm, and the white frames
interpolate between them. Mathematical details of the geodesic
interpolation can be found in \citet{buja2005computational}. The tour
method has been implemented in the R package \CRANpkg{tourr}
\citep{tourr}.

@@

The optimisation problem in the tour context is stated as follows: Given
a randomly generated starting basis \(\mathbf{A}_1\), projection pursuit
finds the final projection basis \(\mathbf{A}_T\) that satisfies the
following optimisation problem:

@@

Below we introduce three line search algorithms in tour:
\texttt{search\_better}, \texttt{search\_better\_random}, and
\texttt{search\_geodesic}. The first two are simulated annealing
algorithms whilst \texttt{search\_geodesic} evaluates neighbourhood
basis to determine the search direction.

\begin{algorithm}
\SetAlgoLined
  \SetKwInOut{input}{input}
  \SetKwInOut{output}{output}
    \input{$f(.)$, $\alpha_1$, $l_{\max}$, $\text{cooling}$} 
    \output{$\mathbf{A}_{l}$}
    generate random start $\mathbf{A}_1$ and set $\mathbf{A}_{\text{cur}} \coloneqq \mathbf{A}_1$, $I_{\text{cur}} = f(\mathbf{A}_{\text{cur}})$, $j = 1$\;
  \Repeat{$\mathbf{A}_l$ is too close to $\mathbf{A}_{\text{cur}}$ in terms of geodesic distance}{
   set $l = 1$\;
  \Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$\;
    update $l = l + 1$\;
  }
  update $\alpha_{j+1} = \alpha_j * \text{cooling}$\;
  construct the geodesic interpolation between $\mathbf{A}_{\text{cur}}$ and $\mathbf{A}_l$\; 
  update $\mathbf{A}_{\text{cur}} = \mathbf{A}_l$ and $j = j + 1$\;
}
  \caption{random search}
  \label{random-search}
\end{algorithm}

\texttt{search\_better} is a random search device that samples a
candidate basis \(\mathbf{A}_{l}\) in the neighbourhood of the current
basis \(\mathbf{A}_{\text{cur}}\) by
\(\mathbf{A}_{l} = (1- \alpha)\mathbf{A}_{\text{cur}} + \alpha \mathbf{A}_{\text{rand}}\)
where \(\alpha\) controls the radius of the sampling neighbourhood and
\(\mathbf{A}_{\text{rand}}\) is a randomly generated matrix with the
same dimension as \(\mathbf{A}_{\text{cur}}\). \(\mathbf{A}_{l}\) is
then orthogonalised to ensure the orthonormal constraint is fulfilled.
When a basis is found with index value higher than the current basis
\(\mathbf{A}_{\text{cur}}\), the search terminates and outputs the basis
for guided tour to construct an interpolation path. The next iteration
of search begins after adjusting \(\alpha\) by a cooling parameter:
\(\alpha_{j+1} = \alpha_j * \text{cooling}\). The termination condition
is when the maximum number of iteration \(l_{\max}\) is reached. The
algorithm of \texttt{search\_better} is summarised in Algorithm
\ref{random-search}. A slightly different cooling scheme has been
proposed by \citet{posse1995projection} to avoid the search space being
reduced too fast. A halving parameter \(c\) is introduced and \(\alpha\)
is only adjusted if the last search takes more than \(c\) times to find
an accepted basis.

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $I_{l} > I_{\text{cur}}$ or $P > U$}{
    generate $\mathbf{A}_{l} = (1- \alpha_j)\mathbf{A}_{\text{cur}} + \alpha_j \mathbf{A}_{\text{rand}}$ and orthogonalise $\mathbf{A}_{l}$\;
    compute $I_{l}  = f(\mathbf{A}_{l})$, $T(l) = \frac{T_0}{\log(l + 1)}$ and $P= \min\left\{\exp\left[-\frac{I_{\text{cur}} -I_{l}}{T(l)}\right],1\right\}$\;
    draw $U$ from a uniform distribution: $U \sim \text{Unif(0, 1)}$\;
    update $l = l + 1$\;
  }
  \caption{simulated annealing}
  \label{simulated_annealing}
\end{algorithm}

Simulated annealing (\texttt{search\_better\_random})
\citep{kirkpatrick1983optimization, bertsimas1993simulated} uses the
same sampling process as \texttt{search\_better} but allows a
probabilistic acceptance of a basis with lower index value based on the
annealing \(T(l)\). Given an initial \(T_0\), the temperature at
iteration \(l\) is defined as \(T(l) = \frac{T_0}{\log(l + 1)}\). When a
candidate basis fails to have an index value larger than the current
basis, simulated annealing gives it a second chance to be accepted with
probability
\[P= \min\left\{\exp\left[-\frac{\mid I_{\text{cur}} - I_{l} \mid}{T(l)}\right],1\right\}\]
where \(I_{(\cdot)}\) denotes the index value of a given basis. This
implementation allows the algorithm to jump out of a local maximum and
enables a more holistic search of the whole parameter space. This
feature is particularly useful when local maxima are present. The
algorithm \ref{simulated_annealing} highlights how simulated annealing
differs from random search in the inner loop.

\begin{algorithm}
\SetAlgoLined
\Repeat{$l > l_{\max}$ or $p_{\text{diff}} > 0.001$}{
  generate $n$ random directions $\mathbf{A}_{\text{rand}}$ \;
  compute $2n$ candidate bases deviate from $\mathbf{A}_{\text{cur}}$ by an angle of $\delta$ while ensure orthogonality\;
  compute the corresponding index value for each candidate bases\;
  determine the search direction as from $\mathbf{A}_{\text{cur}}$ to the candidate bases with the largest index value\;
  determine the step size via optimising the index value on the search direction over a 90 degree window\;
  find the optima $\mathbf{A}_{**}$ and compute $I_{**} = f(\mathbf{A}_{**})$, $p_{\text{diff}} = (I_{**} - I_{\text{cur}})/I_{**}$\;
  update $l = l + 1$\;
}
\caption{search geodesic}
\label{search-geodesic}
\end{algorithm}

\citet{cook1995grand} used a line search algorithm on the space of the
projection bases. In \texttt{search\_geodesic}, the search direction is
computed using the most prominent direction that deviating an tiny angle
of \(\delta\) from the current basis. The step size is chosen by
optimising the index value along the geodesic direction over an 90
degree angle from \(-\pi/4\) to \(\pi/4\) along the search direction
chosen. The optima \(\mathbf{A}_{**}\) is returned for the current
iteration if it meets the percentage improve condition or when
\(l_{\max}\) is reached. Algorithm \ref{search-geodesic} summarises the
inner loop in geodesic search.

@@

\address{%
Nicolas Langrené\\
CSIRO Data61\\
34 Village Street, Docklands VIC 3008 Australia\\
}
\href{mailto:nicolas.langrene@csiro.au}{\nolinkurl{nicolas.langrene@csiro.au}}
